---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Load data

```{r,results="hide"}
activityData <- read.csv("activity.csv")
```

Transform date column from character to Date format

```{r,results="hide"}
activityData$date <- as.Date(activityData$date,'%Y-%m-%d')
```

## What is mean total number of steps taken per day?

Total steps taken per day

```{r}
totalSteps <- tapply(activityData$steps,activityData$date,sum, na.rm=T)
```

A Historgram showing Steps taken each day

```{r}
hist(totalSteps,xlab="Total Steps per day", main="Histogram of Steps per day")
```

Mean & Median of the total number of steps taken per day

```{r}
summary(totalSteps)
```

## What is the average daily activity pattern?

A time series plot of the 5-minute interval and the average number of steps taken (averaged across all days) is shown below:

```{r}
avg_day <- tapply(activityData$steps,activityData$interval,mean,na.rm=T)
plot(avg_day ~ unique(activityData$interval),type="l",xlab="Interval", ylab="Average Steps")
```

5-minute interval, on average across all the days in the dataset that contains the maximum number of steps is:-

```{r}
avg_day[which.max(avg_day)]
```

## Imputing missing values

Total number of missing values in dataset is:-

```{r}
summary(activityData)
```

```{r echo=FALSE}
missingSteps <- sum(is.na(activityData$steps))
```
NA's are in the steps variables. There are `r missingSteps` steps.

Fill the NAs with the average value for that 5-minute interval.

```{r}
newData <- activityData

for (i in 1:nrow(activityData)){
    if(is.na(newData$steps[i])){
        newData$steps[i] <- avg_day[as.character(activityData[i,"interval"])]
    }
    
}
```

Histogram of the total number of steps taken each day

```{r}
newTotalSteps <- tapply(newData$steps,newData$date,sum)
hist(newTotalSteps,xlab="Total steps per day",main="Histogram of Total Steps per Day")
```

Summary of the imputed data set:-

```{r}
summary(newTotalSteps)
```

The imputed data set values differ from the first part of the assignment. Now, the data set values is moving closer to mean values.

## Are there differences in activity patterns between weekdays and weekends?

Add a new column named day

```{r}
newData$day <- ifelse(weekdays(newData$date) %in% c("Saturday","Sunday"),"weekend","weekday")
```

Create subset of data for weekend and weekday for further calculation

```{r}
wendData <- newData[newData$day=="weekend",]
wdayData <- newData[newData$day=="weekday",]
```

Calculate the average steps taken

```{r}
avg_activity_weekend <- tapply(wendData$steps,wendData$interval,mean)
avg_activity_weekday <- tapply(wdayData$steps,wdayData$interval,mean)
```

Transform and combine the subset for weekend and weekday for ploting

```{r}
weekdayData <- data.frame(interval=unique(wdayData$interval),avg=as.numeric(avg_activity_weekday),day=rep("weekday",length(avg_activity_weekday)))
weekendData <- data.frame(interval=unique(wendData$interval),avg=as.numeric(avg_activity_weekend),day=rep("weekend",length(avg_activity_weekend)))

finalData <- rbind(weekdayData,weekendData)
```

Ploting the graph

```{r}
library(lattice)

xyplot(avg ~ interval | day, data = finalData, layout = c(1, 2), type = "l", ylab = "Number of steps")
```

There is differences between weekdays and weekend activity pattern. Weekdays has higher activity between 500 to 800 compare to weekend. Meanwhile weekend has higher activity from 1000 to 1500.